---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# BICA model

<!-- badges: start -->
<!-- badges: end -->

The Bayesian Inseason Chinook Abundance (BICA) model was developed by Aaron Lambert as his Master's thesis project for the University of Alaska Fairbanks CFOS. The project was originally called the "Bayesian Yukon River Canadian-origin Chinook Salmon Inseason Projection Model" and is designed to predict Chinook abundance for the purpose of inseason fishery management of the Alaska Department of Fish & Game (ADF&G). We developed the *bica* package to run the BICA model suite using *CmdStan* and pre-compiled *Stan* code. The benefits of *bica* package are a faster model run time and a more streamlined process.


## Installation

To install *bica*:

  1) Install the R package *CmdStanR* using R commend `install.packages("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))`.
  2) Install R package *instantiate* using R commend `install.packages("instantiate")`
  3) Install *bica* from [GitHub](https://github.com/boppingshoe/bica) with:

``` r
# install.packages("devtools")
devtools::install_github("boppingshoe/bica")
```

## Workflow

1) Have the data set up at a designated directory, then load the data files in the *R* environment:

```{r dir, echo=FALSE}
wd <- "D:/bobby_adfg/projects/yukon_chinook_inseason_mdl/Bayesian-Inseason-Canadian-origin-Chinook-Salmon-Abundance-projection-main"

```


```{r example}
library(bica)

# path for data files
# see Aaron's instructions on types of data needed
dir.data <- file.path(wd, "data")

my_year <- 2024
my_day <- my_day_func(Month = 7, Day = 28)
end_year <- my_year - 1

pf_hist <- readRDS(file.path(dir.data, "Yukon Canadian Chinook PF 29May24.RDS"))
can_hist <- readRDS(file.path(dir.data, "Canadian Chinook Passage RR 2Apr24.RDS"))
pss_hist <-
  readxl::read_xlsx(file.path(dir.data, "ADFG PSS Daily Reports/Yukon Escapement Daily_07292024.xlsx"), skip = 3)
eagle_hist <-
  readxl::read_xlsx(file.path(dir.data, "ADFG Eagle Daily Reports/Yukon Escapement Daily Eagle_07292024.xlsx"), skip = 3)
prior_df_norm <- readRDS(file = file.path(dir.data,"normal curve parameters All Chinook 1995_2023.RDS"))

```

2) Format the data sets into an object using `format_bica_data()`. The object will be used to run BICA model input:

```{r}
bica_data <- bica::format_bica_data(
  my_year, my_day, end_year,
  pf_hist, can_hist, pss_hist, eagle_hist,
  prior_df_norm, prior_df_log = NULL,
  gsi_by_year = NULL, pss_sd = NULL,
  start_day_pss = 148, start_year_pss = 1995
)

```

4) Run BICA model, in this case, model version *pss_prop_es_prop*:

```{r}
model_version <- "pss_prop_es_prop"
n_chains <- 3
n_iter <- 30
n_thin <- 1

bica_out <- bica::run_bica_model(
    bica_data, model_version,
    n_chains, n_iter, n_thin
)

```

5) Summarize results:

```{r}
print(bica_out$stanfit,
      pars = c("alpha", "beta", "sigma", "ln_run_size", "lp__"),
      probs = c(0.1, 0.5, 0.9))

```









