---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# BICA model

<!-- badges: start -->
<!-- badges: end -->

The Bayesian Inseason Chinook Abundance (BICA) model was developed by Aaron Lambert as his Master's thesis project for the University of Alaska Fairbanks CFOS; although he called the project "the Bayesian Yukon River Canadian-origin Chinook Salmon Inseason Projection Model." As the name implied, the BICA model is designed to predict Chinook abundance for the purpose of inseason fishery management of the Alaska Department of Fish & Game (ADF&G). We developed the *bica* package to run the BICA model suite using *CmdStan* and pre-compiled *Stan* code. The benefits of *bica* package are a faster model run time and a more streamlined process.

## Installation

You can install the development version of *bica* from [GitHub](https://github.com/boppingshoe/bica) with:

``` r
# install.packages("devtools")
devtools::install_github("boppingshoe/bica")
```

## Example

Once you have the data set up at a designated directory, you can load the data files in the environment:

```{r dir, echo=FALSE}
wd <- "D:/bobby_adfg/projects/yukon_chinook_inseason_mdl/Bayesian-Inseason-Canadian-origin-Chinook-Salmon-Abundance-projection-main"

```


```{r example}
library(bica)

# path for data files
# see Aaron's instructions on types of data needed
dir.data <- file.path(wd, "data")

my_year <- 2023
my_day <- my_day_func(Month = 8, Day = 9)
end_year <- my_year - 1

pf_hist <- readRDS(file.path(dir.data,"Yukon Canadian Chinook PF 29May24.RDS"))
can_hist <- readRDS(file.path(dir.data,"Canadian Chinook Passage RR 2Apr24.RDS"))
pss_hist <-
  readxl::read_xlsx(file.path(dir.data,"ADFG PSS Daily Reports/Yukon Escapement Daily 9Aug23.xlsx"), skip = 3)
eagle_hist <-
  readxl::read_xlsx(file.path(dir.data,"ADFG Eagle Daily Reports/Yukon Escapement Daily Eagle 9Aug23.xlsx"), skip = 3)

# gsi_by_year <- readRDS(file = file.path(dir.data,"GSI by year unadj 4Apr24.RDS")) # optional

pss_sd <- readRDS(file = file.path(dir.data,"PSS SD 1995_2021.RDS"))
prior_df_log <- read.csv(file = file.path(dir.data,"logistic curve parameters All Chinook 1995_2022.csv"))
prior_df_norm <- readRDS(file = file.path(dir.data,"normal curve parameters All Chinook 1995_2023.RDS"))

```

Compile the input object:

```{r}
bica_data <- format_bica_data(
  my_year, my_day, end_year,
  pf_hist, can_hist, pss_hist, eagle_hist,
  gsi_by_year = NULL,
  pss_sd, prior_df_log, prior_df_norm,
  start_day_pss = 148, start_year_pss = 1995
)

```

Run and summarize results for the BICA model, in this case, model version *PSS prop ES prop*:

```{r}
model_version <- "pss_prop_es_prop"
n_chains <- 4
n_iter <- 50
n_thin <- 1

bica_out <- bica::run_bica_model(
    bica_data, model_version,
    n_chains, n_iter, n_thin
)

bica_out$summary

```










